\renewcommand{\thechapter}{2}

\chapter{Laser cooling and trapping of neutral atoms}

The Nobel Lecture of W.D.Phillips, {\it Laser cooling and trapping of neutral atoms} \cite{phillips1998nobel} tells a great story of the development of laser cooling techniques since the late 1970s. It is recommended for readers who want to learn not only the development of the laser cooling research field but also how researchers approach new physics and technology in a close view. It is aspiring to learn the mature technology that cold atom researcher uses on a daily basis to expand the boundary of physics results from the intelligence, efforts, and persistent pursuit of the early generation of physicists.    

Laser cooling techniques eventually lead to the observation of BEC in 1995, the first Bose-Einstein condensate was created by Eric Cornell, Carl Wieman, and co-workers at JILA on 5 June 1995 \cite{anderson1995observation}. They cooled a dilute vapor of approximately two thousand $^{87} {\rm Rb}$ atoms to below 170~ ${\rm nK}$ using a laser cooling and magnetic evaporative cooling. About four months later, an independent effort led by Wolfgang Ketterle at MIT condensed $^{23} {\rm Na}$ \cite{davis1995bose}. The observation of BEC won laser cooling pioneers Steven Chu, Claude Cohen-Tannoudji, and William D. Phillips the 1997 Nobel Prize in Physics. And Cornell, Wieman, and Ketterle won the 2001 Nobel Prize in Physics for their achievements.

After more than twenty years, the research field of cold atoms is prosperous. The development of laser cooling techniques and the methods to make stable BEC have not only achieved a new state of matter predicted only by Quantum physics, the highly controllable optical and magnetic potential, tunable interaction between particles, and precision measurement techniques have made the BEC system a great platform for Quantum simulation and Quantum computation. 

In this chapter, we use $^{87} {\rm Rb}$ as an example. Start from its energy level, its interaction with light and magnetic field, and then discuss the laser cooling techniques that are essential to make BEC.  

\input{Chapter2_secs/HyperfineStructures}
\input{Chapter2_secs/LaserCoolingTechniques}