%Acknowledgments

\renewcommand{\baselinestretch}{2}
\small\normalsize
\hbox{\ }
 
\vspace{-.65in}

\begin{center}
\large{Acknowledgments} 
\end{center} 

\vspace{1ex}

First of all, I would like to thank my parents and grandparents for everything they have done for me, and their love and support. I began to realize how important they are and how much I miss them after I studied in a foreign country for years. 

Next, I would like to thank my Ph.D. advisor, Dr. Ian Spielman. Ian has been a great advisor. He is always patient in our discussions and always tries to explain physics most intuitively. He would draw nice sketches on the whiteboard which helps him explaining his idea. He would also come to the lab and dig into the hardware together with us and find out why something is not working. When I first joined the group, I had a hard time understanding him because he speaks extremely fast. Over the years, I got much better in communication and I learned a lot from him, not just physics, but also being optimistic when facing uncertainties and difficulties.

During the years, I worked with other members of the RbChip lab. Seiji Sugawa, Francisco Salces-C\'arcoba, Andika Putra, Christopher Billington, and Emine Altuntas, thank you for making RbChip lab an enjoyable place to work. I miss the old times when we can have lunch together, especially after being quarantined at home for half a year in the Covid-19 pandemic.

Lastly, I would like to thank my girlfriend, Dr. Yuchen Yang, who also got her Ph.D. in 2020. We are together nearly the entire time of our Ph.D., we share not only our first name in English but our values of life. I'm very lucky to meet you. 
