\renewcommand{\thechapter}{5}

\chapter{Enhanced transport of SOC Bose gases in disordered potentials: model and simulation}
In materials, microscopic electron scattering processes partly govern the macroscopic conductivity and AL predicts a metal-insulator transition. Increasing a system's conductivity therefore requires some change in these scattering processes.  The most straightforward mechanism is to reduce the disorder strength. Here we describe an alternate approach in which spin-orbit coupling (SOC) greatly suppresses the back scattering and thereby increases the conductivity.  We then propose a realization of this effect using a cold-atom Bose-Einstein condensate (BEC) with laser-induced SOC~\cite{lin2011spin} and disorder from optical speckle.

SOC is a ubiquitous phenomenon in physical systems that describes the interaction between a particle's spin and its momentum. When SOC is combined with a transverse magnetic field (in the sense of Zeeman shifts, not Lorentz forces), gaps in the dispersion relation can open at spin-degeneracy points.  The opening of these gaps modifies the electrons' scattering processes and affects transport. AL was first realized for ultracold-cold atomic systems~\cite{billy2008direct,roati2008anderson}  in 2008, and the experimental techniques are now well established.  Shortly thereafter,  techniques for creating SOC in the cold atom lab were demonstrated~\cite{lin2011spin}.  Together, this makes cold-atom systems an ideal platform to study the interplay between AL and SOC. 

Optical speckle is a powerful tool for creating disordered potentials for atomic systems~\cite{clement2006experimental}.  The strength of the resultant potential is under direct experimental control: the spatial correlation length is tunable and the correlation function is well known.  Here we analytically and numerically study backscattering in speckle potentials of quasi-1d spin-orbit coupled (SOBECs) and compare to the case without SOC. We show that SOC can reduce the scattering processes for specific momentum states. In the broader context, our results suggest that in thin nano-wires, SOC might significantly decrease resistance and improve energy efficiency in electronic devices.

In this chapter, we analytically calculate the probability an initial momentum state being scattered by the speckle potential to any final momentum state and show that SOC can reduce back scattering. Next, we describe numerical simulations of quasi-1d BECs starting in different momentum states subject to a speckle potential with and without SOC. We show that even with the higher order scattering processes and interaction between particles present in the numerical simulations, SOC can reduce the localization effects of disorder and enhance transport.

\input{Chapter5_secs/Scattering}
\input{Chapter5_secs/Simulation}